---
title: "biology 1 EDA my assignment 01"
author: "201900700284 Yang Sihao"
date: "2021/4/19"
output: html_document
---
# hypothesis
null hypothesis: the yeast don't use photoreaction process to repair the pyrimidine dimer mutations.
alternative hypothesis:the yeast use photoreaction process to repair the pyrimidine dimer mutations.
## load the csv
```{r}
y.kill <- read.csv("yeast_uv_kill_curve.csv", stringsAsFactors = T)
```

Create boxplot 
```{r}
boxplot(cell_count ~ dilution, data = y.kill, col = c("red", "orange"))
```

Create a scatter plot
```{r}
plot(cell_count ~ exposure_time, data = y.kill, pch = 20, cex = 1.5, col = dilution)
```

Because of the number of 1/10, so we choose 1/100
```{r}
y.kill.100 <- subset(y.kill, dilution == "1/100")
# plot 
plot(cell_count ~ exposure_time, data = y.kill.100, pch = 20, col = "steelblue", ylim= c(0, 200))
```

We can fit a regression line:
```{r}
y.kill.100.lm <- lm(cell_count ~ exposure_time, data = y.kill.100)
```

Plot adding the mean values for each time point:
```{r}
plot(cell_count ~ exposure_time, data = y.kill.100, pch = 20, col = "steelblue", bty = "n", xlim= c(0, 3), ylim= c(0, 200))
abline(y.kill.100.lm, col = "goldenrod1", lwd = 2)

y.kill.100.mean <- aggregate(cell_count ~ exposure_time, data = y.kill.100, mean)
points(y.kill.100.mean, pch = 18, col = "firebrick", cex = 1.5)
```

We rotate our plot: The x-axis to become the y and the y to become the x-axis.

```{r}
y.kill.100.lm <- lm(exposure_time ~ cell_count, data = y.kill.100)

plot(exposure_time ~ cell_count, data = y.kill.100, pch = 20, col = "steelblue", bty = "n", ylim = c(0, 3), xlim = c(0, 200))
abline(y.kill.100.lm, col = "goldenrod1", lwd = 2)
points(y.kill.100.mean$cell_count, y.kill.100.mean$exposure_time, pch = 18, col = "firebrick", cex = 1.5)
```

Get the LD50:
```{r}
y.kill.100.mid <- (max(y.kill.100.mean) - min(y.kill.100.mean)) / 2

y.kill.100.ld50 <- predict(y.kill.100.lm, data.frame(cell_count = y.kill.100.mid))
paste("UV-C LD50 exposure time is:", round(y.kill.100.ld50, 1))
```

# Data analysis

Read the experiment data:
```{r}
rm(list = ls())
y <- read.csv("yeast_uv_irradiation.csv", stringsAsFactors = T)
```

ANOVE and TurkeyHSD
```{r}
y.aov <- aov(cell_count ~ treatment, data = y)
summary(y.aov)
TukeyHSD(y.aov)
```
There is look no difference, should NOT jump in data anlysis

```{r}
boxplot(cell_count ~ treatment, data = y)
```

We have extreme outliers in our data
```{r}
palette(c("blue", "red", "darkgreen"))
plot(cell_count ~ plate, data = y, col = treatment, pch = 20, cex = 1.5)
legend("topright", levels(y$treatment), pch = 20, col = 1:nlevels(y$treatment))
```

We have no other choice but to remove these outliers from our analysis.
```{r}
y.fix <- subset(y, cell_count < 250)
```

Create a boxplot:
```{r}
boxplot(cell_count ~ treatment, data = y.fix)
```

Show the scatter plot:
```{r}
plot(cell_count ~ plate, data = y.fix, col = treatment, pch = 20, cex = 1.5)
legend("topright", levels(y$treatment), pch = 20, col = 1:nlevels(y$treatment))
```

The statisitical analysis:
```{r}
y.fix.aov <- aov(cell_count ~ treatment, data = y.fix)
summary(y.fix.aov)
TukeyHSD(y.fix.aov)
```

```{r}
plot(TukeyHSD(y.fix.aov), col = "red")
```


