---
title: "EDA biology assignment02"
author: "201900700284 Yang Sihao"
date: "2021/5/11"
output: html_document
---

# Load the data
```{r}
data <- read.csv("seeds.csv", stringsAsFactors = T)
data
```

 load the samples
```{r}
samples <- read.csv("sample.csv", stringsAsFactors = T)
samples
```

```{r}
str(samples)
```

```{r}
samples$pressure<-factor(samples$pressure,c("0","-0.5","-1"))
str(samples)
```

```{r}
seeds<-merge(samples,data)
seeds
```

remove the unnecessary data.
```{r}
rm(data, samples)
```

# Calculate the Mean Time to Germination(MTG)
```{r}
seeds$interval_3  <- seeds$total_3
seeds$interval_5  <- seeds$total_5  - seeds$total_3
seeds$interval_7  <- seeds$total_7  - seeds$total_5
seeds$interval_10 <- seeds$total_10 - seeds$total_7
seeds$interval_12 <- seeds$total_12 - seeds$total_10
seeds$interval_14 <- seeds$total_14 - seeds$total_12

seeds
```

Create a new function to calculate the weighted mean using the time periods in the experiment.
```{r}
calcMTG <- function(weights){
  weighted.mean(c(3, 5, 7, 10, 12, 14), weights)
}
```

```{r}
calcMTG(c(0, 0, 0, 0, 1, 1))
```

```{r}
calcMTG(c(1, 0, 1, 0, 0, 2))
```

```{r}
mtg <- apply(seeds[, c("interval_3", "interval_5", "interval_7", "interval_10", "interval_12", "interval_14")], 1, calcMTG)
mtg
```

Add the calculated MTG values as a new column to the data.frame.
```{r}
seeds$MTG <- round(mtg, 2)
rm(mtg)

seeds
```


# Calculate the Germination Fraction (GF)
```{r}
seeds$GF <- seeds$total_14 / 5
seeds
```

# Perform ANOVA on the MTG 
ANOVA model 
```{r}
seeds.mtg.aov <- aov(MTG ~ species * pressure + shelf, data = seeds)
```

ANOVA table
```{r}
summary(seeds.mtg.aov)
```

Diagnostic plots:
```{r}
par(mfrow = c(2, 2))
plot(seeds.mtg.aov)
```

```{r}
par(mfrow=c(1,1))
```

Means in the groups:
```{r}
seeds.mtg.means <- model.tables(seeds.mtg.aov, "means")
seeds.mtg.means$tables$`species:pressure`
```

TukeyHSD pairwise tests:
```{r}
seeds.mtg.hsd <- TukeyHSD(seeds.mtg.aov)

par(mar = c(5, 12, 2, 2))
par(mfrow = c(2, 2))
plot(seeds.mtg.hsd, las = 1, col = "red")
```

```{r}
par(mfrow = c(1, 1))
```

Interaction plot:
```{r}
par(mar = c(5, 5, 2, 2))
interaction.plot(seeds$pressure, seeds$species, seeds$MTG, col = c("darkolivegreen4", "goldenrod3"), type = "b", lty = 1, lwd = 3, pch = c(20, 20), trace.label = "species", xlab = "pressure", ylab = "MTG")
```

# Graphical presnetation of MTG
```{r}
par(mar = c(10, 5, 2, 2))
boxplot(MTG ~ species + pressure, data = seeds, las = 2, col = c("darkolivegreen", "gold4", "darkolivegreen4", "goldenrod3", "darkolivegreen2", "gold"), frame = F, outline = F, sep = ": ", xlab = "")
stripchart(MTG ~ species + pressure, data = seeds, vertical = T, method = "jitter", jitter = 0.15, pch = 20, col = "darkgreen", add = T)
```

# Perform ANOVA on the GF
ANOVA model
```{r}
seeds.mtg.aov <- aov(GF ~ species * pressure + shelf, data = seeds)
```

ANOVA table
```{r}
summary(seeds.mtg.aov)
```

Diagnostic plots:
```{r}
par(mfrow = c(2, 2))
plot(seeds.mtg.aov)
```

```{r}
par(mfrow = c(1, 1))
```

Means in the groups:
```{r}
seeds.mtg.means <- model.tables(seeds.mtg.aov, "means")
seeds.mtg.means$tables$`species:pressure`
```

TukeyHSD pairwise tests:
```{r}
seeds.mtg.hsd <- TukeyHSD(seeds.mtg.aov)

par(mar = c(5, 12, 2, 2))
par(mfrow = c(2, 2))
plot(seeds.mtg.hsd, las = 1, col = "red")
```

```{r}
par(mar = c(5, 5, 2, 2))
interaction.plot(seeds$pressure, seeds$species, seeds$GF, col = c("darkolivegreen4", "goldenrod3"), type = "b", lty = 1, lwd = 3, pch = c(20, 20), trace.label = "species", xlab = "pressure", ylab = "GF")
```

# Graphical presentation of GF
```{r}
par(mar = c(10, 5, 2, 2))
boxplot(GF ~ species + pressure, data = seeds, las = 2, col = c("darkolivegreen", "gold4", "darkolivegreen4", "goldenrod3", "darkolivegreen2", "gold"), frame = F, outline = F, sep = ": ", xlab = "")
stripchart(GF ~ species + pressure, data = seeds, vertical = T, method = "jitter", jitter = 0.15, pch = 20, col = "darkgreen", add = T)
```

