---
title: "Assignment2"
author: "U7457462_Sihao_Yang"
date: "2022-10-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# i. Statistical Analysis and Interpretation (50%)
## 1. Correct analysis of Clark et al. (2020) data to generate the summary statistics (means, SD, N) for each of the fish species’ average activity for each treatment.
### Libraries and Packages

Load in the corresponding libraries, tidyverse as well as readxl.
```{r}
library(tidyverse)
library(readr)
install.packages("pacman", repos = "http://cran.us.r-project.org")
library(pacman)
p_load(bookdown, tidyverse, ggforce, flextable, latex2exp, png, magick) # list all the packages 
```

### Import CSV files
The data is in an Excel file (“OA_activitydat_20190302_BIOL3207.csv”) 
```{r}
data_Clark_path <- "data/OA_activitydat_20190302_BIOL3207.csv"
data_Clark <- read_csv(data_Clark_path)
data_Clark
```

First, write some code below to remove missing data
```{r}
# Code to removing missing data from the `OA_activitydat_20190302_BIOL3207.csv` data frame. 
data_Clark <- na.omit(data_Clark)
data_Clark
```
### Create the table of mean/SD/N of data
```{r}
# Use flextable to render the summary table in a tidy format
mean_sd <- list(mean=~mean(.x, na.rm = TRUE), 
  sd = ~sd(.x, na.rm = TRUE)
)
data_Clark %>% select(species,treatment,sl,size,activity) %>% group_by(species,treatment,size)%>% summarise(across(where(is.numeric), mean_sd),N=n()) %>% flextable(,col_keys = c("species","treatement","size","sl_mean","sl_sd","activity_mean","activity_sd","N")) %>% add_header_row(, values = c("", "sl","activity",""), colwidths = c(3, 2,2,1)) %>% align(, i = 1, part = "header", align = "center")
```

## 2.Through coding, merge the summary statistics generated from 1) with the metadata (i.e., clark_paper_data.csv) from Clark et al. (2020).






